---
title: "Exercise 1"
author: "Andrew Boscence and Carlos Munoz - ID7"
date: "February 22nd  2016"
output: html_document
---

### Introduction

Wikipedia is a free encyclopedia written collaboratively by volunteers around the world. A small part of Wikipedia contributors are administrators, who are users with access to additional technical features that aid in maintenance. In order for a user to become an administrator a Request for adminship (RfA) is issued and the Wikipedia community via a public discussion or a vote decides who to promote to adminship. Using a complete dump of Wikipedia page edit history (from January 3 2008) we extracted all administrator elections and vote history data. This gave us nearly 2,800 elections with around 100,000 total votes and about 7,000 users participating in the elections (either casting a vote or being voted on). Out of these 1,200 elections resulted in a successful promotion, while about 1,500 elections did not result in the promotion. About half of the votes in the dataset are by existing admins, while the other half comes from ordinary Wikipedia users.

### Exercise 1

In this exercise, you will learn how to use the Igraph package, a powerful tool to analyze and visualize networks in R. The main goal in this exercise is to acquire programming abilities to extract value from relational data. The exercise is divided in 10 sections. You will receive a grade from 0 to ten. The exercise can be done individually or in pairs. Please, send an email with the members of the pair to allorente@faculty.ie.edu .

To find more information about R markdown, see http://rmarkdown.rstudio.com/articles.html.

<span style="color:blue">
The answers in every section, please code it in blue
</span>

### Load data

We have two datasets: one for loading metadata about elections and the other to load the adjacency list. Columns in votings dataset are: "voter"-"user"-"timestamp"-"vote". 
In this section you will show the ten first lines of every loaded dataset. 

```{r}
#loading some libraries
library(igraph)
library(ggplot2)
library(dplyr)
library(plotrix)
```

```{r}
setwd("~/2.IE/Term 2 - Winter/Social Network Analysis/Exercise 1")
# load raw data files
metadata<-read.csv("metadata.txt",sep=";",stringsAsFactors = F,header=F,quote="")
# setting headers
colnames(metadata)<-c("user","nominator","deadline","result")
head(metadata,10)
```

```{r}
# load raw data files
votings<-read.csv("votings.txt",sep=";",stringsAsFactors = F,header=F,quote="")
# setting headers
colnames(votings)<-c("user","candidate","datetime","vote")
head(votings,10)
```


### Build the network object

On this chunk we create the graph object by using the function graph.edgelist and then we set the attributes "vote" and "timestamp" to the edges. Maybe this piece of documentation is useful for this purpose http://www.inside-r.org/packages/cran/igraph/docs/attributes. After creating the graph object, use the summary function to generate the following output:

```{r}
#CHUNK 2: generate the graph object and visualize the summary
#transform into graph
g<-as.matrix(votings)
g<-graph_from_edgelist(g[,1:2])
g<-set.edge.attribute(g,"vote",value=votings[,4])
g<-set.edge.attribute(g,"timestamp",value=votings[,3])

```

```{r}
summary(g)
```

What do all the elements in the output mean?
<span style="color:blue">
"Summary" is telling me:
- D --> the graph is Directed
- N --> the graph is Named (it has the "name" vertex attribute set)
- 7194 --> the first number is the number of vertices in the graph
- 114040 --> the second number is the number of edges in the graph
- "name (v/c)" --> indicates that "name" is a vertex attribute of type character 
- "vote (e/n)" --> indicates that "vote" is an edge attribute of type numeric
- "timestamp (e/c)" --> indicates that "timestamp" is an edge attribute of type character
</span>


### Degree distribution

You have to plot here the degree distribution of the nodes in the network, a typical analysis to visually understand the heterogeneity of connections within the network. You have to analyze the node degree distribution in three different ways: general degree, indegree and outdegree. To do this, use the degree function (http://www.inside-r.org/packages/cran/igraph/docs/degree) and the plot function. Consider using logarithmic scales in the plot function to get a better visualization. 

```{r}
#CHUNK 3: compute the three types of degree methods and plot the distribution
degree_in = degree(g,v = V(g), mode = "in")
plot(density(degree_in),log="xy", main = "Density Plot for the In Degree",xlab = "In Degree",col = "blue")
points(density(degree_in),col = "blue")

degree_out = degree(g,v = V(g), mode = "out")
plot(density(degree_out),log="xy", main = "Density Plot for the Out Degree",xlab = "Out Degree",col = "blue")
points(density(degree_out),col = "blue")

degree_total = degree(g,v = V(g), mode = "total")
plot(density(degree_total),log="xy", main = "Density Plot for the Total Degree",xlab = "In Degree",col = "blue")
points(density(degree_total),col = "blue")


```

<span style="color:blue">
After analyzing the results, please answer the next question: Is it a scale-free network? Consider using the function power.law.fit to provide a statistically robust answer.

A power-law distribution is a distribution where the frequency of an event varies as a power of that event. In this excercise, this means that the degree frequency varies as a power of the degree.

To test this, we set the null hypothesis so that the distribution is equal to a power-law distribution. The alternative is that the distribution is not equal to a power-law distribution. We now run the power.law.fit function on all three degree types to test the hypothesis.
</span>

```{r}
#CHUNK 4: show statistically whether the distribution follows a power-law function.
pl_in = power.law.fit(degree_in)
pl_in$KS.p
pl_out = power.law.fit(degree_out)
pl_out$KS.p
pl_total = power.law.fit(degree_total)
pl_total$KS.p
```

<span style="color:blue">
It is obvious that this is not a random network. However, we also see that the KS.P Value for all three degree types is above the .05 threshhold. Therefore we fail to reject the null hypothesis that the distribution is equal to a power-law distribution

https://www.stat.berkeley.edu/~aldous/Research/Ugrad/Willy_Lai.pdf
https://www.youtube.com/watch?v=dNbSWsQGHsw
</span>


### Reciprocity

Search for the function to compute reciprocity in Igraph package. How many reciprocal links are there in this network? Analyze the results: Is reprocity a good measure to analyze this network?

```{r}
#CHUNK 5: compute reciprocity on voting network
reciprocity(g)
```

<span style="color:blue">
Reciprocity measures the proportion of mutual connections in a directed graph. 
I think it is not a good measure to analyze the network if there are many administrators with low voting activity, that have casted a vote only once or twice and are not linked to other users. Perhaps Reciprocity is more useful in a subgraph with more connected components.
</span>

### Connected components

In this section we want to compute the connected components of the graph, that is, the subgraphs in which any two vertices are connected to each other by paths. Analyze what happens if directions of links are considered or not and explain why. You have to use the clusters function to get the connected components.

<span style="color:blue">
Connected graphs are graphs where you can get to any node from any other node. If you use direction, they can be either strongly connected or weakly connected. Weakly connected graphs come about when direction dictates the way you must go from one node to another, they have restricted paths. Strongly connected graphs mean you can take any route. 
</span>

```{r}
#CHUNK 6: use the clusters function to get the connected components and analyze their size.
clust<-clusters(g)
#clust$membership
#I want to view how many clusters of each size I have
table(clust$csize)

```

<span style="color:blue">
"clust" is storing the information on the connected components of the graph. 
By doing a table out of clust$size we obtain how many clusters of each size are present in the graph
</span>

We want to keep the main connected component in a different variable. To do this, you have to use the function induced.subgraph and use the result of the clusters function.

```{r}
#CHUNK 7: create a new graph object g2 modelling the main connected component and show the summary of this object.
g2<-induced.subgraph(g,clust$membership==1)

#visual check
summary(g2)
```

<span style="color:blue">
Note: we have named the main connected component"subg" instead of "g2"
In "summary" I check that the resulting graph contains the main connected component, which in this case is composed of 7136 out of the 7194 original vertices
</span>


### Shortest paths

Given two nodes, the shortest path betweem them is the subset of links connecting both nodes going from one node to another adjacent one. Two typical measures to analyze a graph is the average length of the shortest paths or the maximum one (diameter). Use the shortest.paths function to compute both average path length and diameter on the original graph. What value do you get? Can you explain why? Now compute both metrics for the main connected component of the previous section. Is it a small-world network? Explain similarities and differences.

```{r}
#CHUNK 8: compute Avg. path length and diameter for both the original graph and the main connected component.
#storing distance among nodes for original graph, then calculating diamenter and avg path
short_g<-shortest.paths(g)
max(short_g)
mean(short_g)

#storing distance among nodes for the main connected component, then calculating diameter and avg path
short_subg<-shortest.paths(g2)
max(short_subg)
mean(short_subg)

#growing a random network of same edge and vertex value as the main connected component. Calculating diameter and avg path
randomg<-erdos.renyi.game(n=vcount(g2),p.or.m=ecount(g2),type=c("gnm"),directed=T,loops=F)
summary(randomg)                          #checking the graph is also D, with same number of vertices and edges
short_randomg<-shortest.paths(randomg)
max(short_randomg)
mean(short_randomg)

#calculating clustering coefficients of both the principal component and the ramdom graph
transitivity(g2)
transitivity(randomg)

```

<span style="color:blue">
As the original graph is not a connected component (it contains 31 clusters of different sizes) it makes sense to have both the mean and the max distance as infinite: some nodes are not connected among themselves.
On the other hand all nodes in the main connected component are connected among themselves, therefore the mean and max provides a finite value.
For analizing if the network is a small-world one, I have to analyze two things:
- Transitivity: A high transitivity indicates small-worldness
- Mean shortest path: a low mean shortest path indicates also small-worldness: many nodes are few nodes away from each others
The transitivity of the connected component is much higher than the transitivity of the random graph, while the mean is in the same range. I would say it is a small network.
However, this is only if we limit our scope to the main connected component. Otherwise, our entire network is not small world because parts of it are unreachable.
</span>

### Centrality

There are many ways of computing centrality. In this section we will compare two of the most widely used methods: degree and betweenness. Ignore directions of edges in this section. What are the theoretical differences comparing these two methods? Analyze the differences based on the results.

```{r}
#CHUNK 9: compute degree and betweenness per node in order to analyze the differences between these two algorithms.

#computing degree and betweenness for the main connected component
between_g = betweenness(g, v=V(g), directed = F, weights = NULL,nobigint = F, normalized = FALSE)
degree_g = degree(g)

#plotting using logarithmic scales
plot(degree_g,between_g, log = "xy", main = "Plot of Degree and Betweenness for Wikipedia voting network"
     ,xlab = "Degree (All)", ylab = "Betweenness", col = "blue")
```

<span style="color:blue">
The plot of Degree and Betweenness shows that the betweenness differs greatly from nodes with low degrees. This indicates that some nodes with lower degrees may be considered more central when using the betweenness method. Also, some nodes considered to be moderately central with the betweenness method can vary in ther centrality using degree. However, we notice that nodes with the highest centrality in one method tend to also be considered central in the other method. We see this as the chart begins to thin when we move to the top right.   
</span>

### Introduction to graph visualization

In this final section we will explore how to visualize networks in R. We are going to use a layout, layout.lgl, to place the nodes and we are going to colour them depending on they were i) candidates and accepted as administrators (green) ii) candidates and weren't accepted as administrator (red) iii) nominators (blue) or iv) only voters (yellow). 

```{r}
#CHUNK 10: use layout.lgl as layout and ifelse sentences to set the colour vector, then use the plot function to finally plot the graph.
#Getting the winners of the elections
winners<-metadata%>%
  filter(result==1)

#Adding a "color" attibute to vertices
cols<-as.vector(rep("red",times=vcount(g2)))
set.vertex.attribute(g2,"color",value=cols)

#assignning colours according to specifications
V(g2)$color<-ifelse(V(g2)$name %in% winners$user,"green",ifelse(V(g2)$name %in% metadata$user,"red",ifelse(V(g2)$name %in% metadata$nominator,"blue","yellow")))

#plottinh a graph with the proper settings for good visualization
plot.igraph(g2, layout=layout.lgl, main="Wikipedia Election Votes",vertex.size=3,edge.arrow.size=0.01, vertex.label="")

legend("topleft",legend=c("Winners","Losers", "Nominators", "User"), col=c("Green","Red","blue", "yellow"), pch=19,title="Network Legend")
```

<span style="color:blue">
The visualization will be performed on the main connected component, which holds the vast majority of administrators.
This visualization arranges vertices in a circle, where the position of each vertex depends on its degree: the higher the degree, the closer to the center.
A visual inspection indicates that nominees tend to be well connected with other administrators, with almost all of them in the inner part of the circle. Also, it seems that the likelihood of winning an election is directly related with proximity to the  center of the graph: better connections often means more positive votes.
A good network is required for winning an election. This means that points further from the center will have a lower chance of being elected. It is worth noting that all the poorly connected candidates (outer elements of the circle) have lost their elections.
There seems to be a large overlap with nominators and nominees: there are very few blue vertices, and with 1.200 elections held that would signal that many of the nominators have also been nominated to administrater a page (so they appear in red and green).
The graph also tells a lot about administrators' activity. There seems to be two large groups: very active groups (inner circle, nominating, voting and being nominated to administer several pages) and passive group (outer section, not nominating or being nominated and merely cast votes).
</span>